---
title: "Non linear fitting \n(worksheet)"
author: "Etienne Low-Décarie"
date: '2018-01-31'
output: 
  html_document: 
    toc: yes
---

#Housekeeping

Load required and suggested packages
```{r echo=FALSE, message=FALSE}
require(ggplot2)
require(cowplot)
require(dplyr)
```

# Examples

## Exponential growth

\[N_t = N_0e^{(rt)}\]

```{r echo=FALSE}
exp_growth <- function(N0=10,
                       r=1.1,
                       t=0:10){
  N = N0 * exp(r*t)
}

p <- qplot(x=0:10, 
      y=exp_growth(t=0:10),
      xlab="Time",
      ylab="N")+
  geom_line()

print(p)
```

## Logistic growth

\[\frac{dN}{dt} = rN(\frac{K-N}{K})\]


\[N_t = \frac{K}{1+(\frac{K}{N_0}-1)e^{(-rt)}}\]

```{r echo=FALSE}
logistic_growth <- function(N0=10,
                       r=1.1,
                       K=4000,
                       t=0:10){
  N = K/(1+(K/N0-1)*exp(-r*t))
}

p <- qplot(x=0:15,
      y=logistic_growth(t=0:15),
      xlab="Time",
      ylab="N")+
  geom_line()

print(p)
```

## Rate of growth in a logistic equation

```{r echo=FALSE}
p <- qplot(x=logistic_growth(t=0:15),
      y=lead(logistic_growth(t=0:15))-logistic_growth(t=0:15),
      xlab="N(t)",
      ylab="N(t+1)-N(t)")+
  geom_line()

print(p)
```


Quadratic equation

\[\Delta N = r(K-\alpha N)N\]

```{r echo=FALSE}
logistic_rate_growth <- function(N0=10,
                                 K=4000,
                       r=1.1,
                       alpha=0.1,
                       N=0:4000){
  delta_N = r*N-r*alpha*N^2
}

p <- qplot(x=0:4000,
      y=logistic_rate_growth(N=0:4000),
      xlab="N",
      ylab="delta_N")+
  geom_line()

print(p)
```

## Thermal response




# Simulating data

## Why simulated data

## Writing a function

### Exponential growth

```{r}
exp_growth <- function(N0=100,
                       lambda=1.5,
                       t=0:10){
  N = N0 * lambda^t
}
```

## Adding noise

```{r}
set.seed(20)
```


# Fitting the model to data

## Linearizable models

### Exponential growth

```{r}
p <- qplot(x=0:10, 
      y=log(exp_growth(t=0:10)),
      xlab="Time",
      ylab="log(N)")+
  geom_line()

print(p)
```

## Logistic growth

Does not work for logistic

```{r echo=FALSE}

p1 <- qplot(x=0:15,
      y=log(logistic_growth(t=0:15)),
      xlab="Time",
      ylab="log(N)")+
  geom_line()


p2 <- qplot(x=log(0:15),
      y=log(logistic_growth(t=0:15)),
      xlab="log(Time)",
      ylab="log(N)")+
  geom_line()

p3 <- qplot(x=log(0:15),
      y=logistic_growth(t=0:15),
      xlab="log(Time)",
      ylab="N")+
  geom_line()

plot_grid(p1)
```



##  Using linear fitting for polynomials

```{r}
delta_N <- data.frame(Nt=logistic_growth(t=0:15),                      delta_N=lead(logistic_growth(t=0:15))-logistic_growth(t=0:15))

r <- 1.1
K <- 4000

fit <- lm(delta_N~Nt+I(Nt^2),
          data=delta_N)


```



addition, subtraction, multiplication, and non-negative integer exponents of variables. 

https://stackoverflow.com/questions/17457884/using-lmpoly-to-get-formula-coeff

## Non linear least squares

## Likelihood approach

## 

# Starting values

## Manual

## Self start models

# Robust fitting

# Assessing error

What is the confidence interval around each parameter?

## Error across replicates

## Boostrapping

## Model sensitivity

### Variance-covariance and MCMC



# Next steps (or what we won't cover)

## Mixed effect non-linear models

## Multi-demensional and equation systems

## Bayesian approaches

# Bestiary of packages

# References and ressources

http://rstudio-pubs-static.s3.amazonaws.com/28730_850cac53898b45da8050f7f622d48927.html

https://rstudio-pubs-static.s3.amazonaws.com/252603_9c6a71110dc74cc7832f154449235f7f.html

https://www.r-bloggers.com/learning-r-parameter-fitting-for-models-involving-differential-equations/